Automatic Matching of Bullet Lands

### 2016-05-07 revision/response.Rmd

9. Would it be possible to stress the classification system by only using part of the profile? Such as would occur when a partial bullet is recovered? By stressing the random forest you would be able to compare it to other classifiers.

***Yes, it would be possible to stress the classification system by using only parts of the profile, but a systematic evaluation of this by far exceeds the scope of this paper.***

10. What is the standard error of the error rates presented for the matching algorithms? The number of cross comparisons gives false impression the standard errors the when presenting the results.

***Under the assumption of land-to-land independence, the number of land-to-land matches is distributed according to a Binomial(N, p) distribution where N is the total number of bullet lands and p is the probability of a match between two bullet lands. However, the bullet lands from the Hamby study are all known to be either pairwise matches or non-matches. Thus, we present the Out-of-bag (OOB) error for the random forest, representing the error rate as determined from the observations not present in the particular bootstrap samples. This is discussed further on Page 19.***

### 2016-05-07 revision/response.Rmd

21. p. 3, l. -10:  Be explicit that grooves and striations are synonyms.

***Grooves and striations are actually not synonyms in this case. The grooves are impressions induced by the rifling of the barrel, and properties associated with them are class characteristics not unique to individual gun barrels. We've taken care to ensure the distinction is clear between the striae of interest, which are extracted from the lands of the bullet (between grooves), and the grooves themselves.***

22. Early on in the paper, it might be helpful to mention the typical width of a striation, the number of striations typically seen on a fired bullet, and so forth.

***We have added the typical width of both the groove-to-groove (land) width, as well as the typical width of a striation.***

### 2016-05-07 imaging-paper.Rnw
References for striae width

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\textwidth}\centering
\caption{View of the data along the circumference of the bullet (circular segment of about 30 degrees).\label{fig:sidex3p}}{%
      \includegraphics[width=.65\textwidth]{images/sidex3p.png}
      \hspace{1cm}
      \includegraphics[width=.25\textwidth]{images/side-sketch.png}
    }
\end{subfigure}    
\begin{subfigure}[t]{\textwidth}\centering
    \caption{Frontal view of a bullet land (lower end of the view is the bottom of the bullet). \label{fig:topx3p}}{%
    \includegraphics[width=.65\textwidth]{images/topx3p.png} \hspace{1cm}
    \includegraphics[width=.15\textwidth]{images/top-sketch.png}
    }
\end{subfigure}
\caption{Example of a groove-to-groove scan of a bullet land. The red-dotted rectangle on the right shows the location and orientation of the segment. }
\end{figure}
Each fired bullet is provided in the form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land". For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number. An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}. These figures show side and top profiles of the land respectively. The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artifact, but a direct and expected consequence of the spin induced by the rifling during the firing process. Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt. The direction of the rifling is a class characteristic, i.e.\ a feature that pertains to a particular class of firearms, and is not unique at the individual barrel/bullet level.

>>>>

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\textwidth}\centering
\caption{View of the data along the circumference of the bullet (circular segment of about 30 degrees).\label{fig:sidex3p}}{%
      \includegraphics[width=.65\textwidth]{images/sidex3p.png}
      \hspace{1cm}
      \includegraphics[width=.25\textwidth]{images/side-sketch.png}
    }
\end{subfigure}    
\begin{subfigure}[t]{\textwidth}\centering
    \caption{Frontal view of a bullet land (lower end of the view is the bottom of the bullet). \label{fig:topx3p}}{%
    \includegraphics[width=.65\textwidth]{images/topx3p.png} \hspace{1cm}
    \includegraphics[width=.15\textwidth]{images/top-sketch.png}
    }
\end{subfigure}
\caption{Example of a groove-to-groove scan of a bullet land. The red-dotted rectangle on the right shows the location and orientation of the segment. }
\end{figure}

<<width, echo=FALSE>>=
typical_width <- read.csv("csvs/grooves.csv") %>%
    summarise(length = mean(groove_right - groove_left)) %>%
    as.numeric %>%
    round(digits = 2)
@

Each fired bullet is provided in the form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land". In the Hamby data, typical length (groove-to-groove) of a land is about \Sexpr{typical_width} micrometers or \Sexpr{typical_width / 1000} millimeters. For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number. An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}. These figures show side and top profiles of the land respectively. The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artifact, but a direct and expected consequence of the spin induced by the rifling during the firing process. Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt. The direction of the rifling is a class characteristic, i.e.\ a feature that pertains to a particular class of firearms, and is not unique at the individual barrel/bullet level.

The typical number and width of striation markings on bullets varies significantly depending on the gun barrel. For instance, a Smith and Wesson barrel with a land-width of 2.4 millimeters contained an average 60 striae, with an average width of about 0.08 millimeters~\citep{chu:2011}.

### 2016-05-07 imaging-paper.tex
added boxplots - densities do not work, because they are too skew
References for striae width

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\textwidth}\centering
\caption{View of the data along the circumference of the bullet (circular segment of about 30 degrees).\label{fig:sidex3p}}{%
      \includegraphics[width=.65\textwidth]{images/sidex3p.png}
      \hspace{1cm}
      \includegraphics[width=.25\textwidth]{images/side-sketch.png}
    }
\end{subfigure}    
\begin{subfigure}[t]{\textwidth}\centering
    \caption{Frontal view of a bullet land (lower end of the view is the bottom of the bullet). \label{fig:topx3p}}{%
    \includegraphics[width=.65\textwidth]{images/topx3p.png} \hspace{1cm}
    \includegraphics[width=.15\textwidth]{images/top-sketch.png}
    }
\end{subfigure}
\caption{Example of a groove-to-groove scan of a bullet land. The red-dotted rectangle on the right shows the location and orientation of the segment. }
\end{figure}
Each fired bullet is provided in the form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land". For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number. An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}. These figures show side and top profiles of the land respectively. The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artifact, but a direct and expected consequence of the spin induced by the rifling during the firing process. Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt. The direction of the rifling is a class characteristic, i.e.\ a feature that pertains to a particular class of firearms, and is not unique at the individual barrel/bullet level.

The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match. 
The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned. 
% Hamby study: automated systems were successfully used for matching:
%  Intelligent Automation's SciClops - Dr. Ben Bachrach (Maryland, United States);
% Automated Land Identification System (ALIS) - Mr. Tsuneo Uchiyama (Tokyo, Japan);
% Integrated Ballistics Identification System (IBIS) – Mr. Robert Thompson (California, United States);
% BulletTRAX-3DTM - Forensic Technology Scientists (Montreal, Canada) (2 sets)

@@

>>>>

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\textwidth}\centering
\caption{View of the data along the circumference of the bullet (circular segment of about 30 degrees).\label{fig:sidex3p}}{%
      \includegraphics[width=.65\textwidth]{images/sidex3p.png}
      \hspace{1cm}
      \includegraphics[width=.25\textwidth]{images/side-sketch.png}
    }
\end{subfigure}    
\begin{subfigure}[t]{\textwidth}\centering
    \caption{Frontal view of a bullet land (lower end of the view is the bottom of the bullet). \label{fig:topx3p}}{%
    \includegraphics[width=.65\textwidth]{images/topx3p.png} \hspace{1cm}
    \includegraphics[width=.15\textwidth]{images/top-sketch.png}
    }
\end{subfigure}
\caption{Example of a groove-to-groove scan of a bullet land. The red-dotted rectangle on the right shows the location and orientation of the segment. }
\end{figure}

Each fired bullet is provided in the form of a set of six x3p files, where each file is a surface scan between adjacent grooves on the bullet, called a ``land". In the Hamby data, typical length (groove-to-groove) of a land is about 2005.9 micrometers or 2.01 millimeters. For notational simplicity, we refer to a particular land of some bullet as bullet X-Y, where X is the bullet identifier, and Y is the land number. An example of plotting one of these lands is given in Figures~\ref{fig:sidex3p} and~\ref{fig:topx3p}. These figures show side and top profiles of the land respectively. The tilt of the lines to the left in Figure~\ref{fig:topx3p} is not an artifact, but a direct and expected consequence of the spin induced by the rifling during the firing process. Depending on whether a barrel is rifled clockwise or counter-clockwise, the striations have a left or right tilt. The direction of the rifling is a class characteristic, i.e.\ a feature that pertains to a particular class of firearms, and is not unique at the individual barrel/bullet level.

The typical number and width of striation markings on bullet varies significantly depending on the gun barrel. For instance, a Smith and Wesson barrel with a land-width of 2.4 millimeters contained an average 60 striae, with an average width of about 0.08 millimeters~\citep{chu:2011}.

The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match. 
The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate, no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned. 
% Hamby study: automated systems were successfully used for matching:
%  Intelligent Automation's SciClops - Dr. Ben Bachrach (Maryland, United States);
% Automated Land Identification System (ALIS) - Mr. Tsuneo Uchiyama (Tokyo, Japan);
% Integrated Ballistics Identification System (IBIS) – Mr. Robert Thompson (California, United States);
% BulletTRAX-3DTM - Forensic Technology Scientists (Montreal, Canada) (2 sets)

@@

\includegraphics[width=.6\textwidth]{figure/tree:forest-2}
